---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 1"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Peter Meihuizen"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch Univerity, South Africa" # First Author's Affiliation
Email1: "21831041\\@sun.az.za" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
 # Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

For this question I am looking at how different continents were affected by Covid-19. 

# Analysis \label{Analysis}

Ok so firstly I need to import the data.

```{r}
library(csvread)
coviddf <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_1\\data\\Covid\\owid-covid-data.csv")
source("code/create_covid_plots.R")
source("code/data_transformation.R")
source("code/smoke_data_trans.R")
source("code/resp_trans.R")
source("code/merge_resp_hosp.R")
source("code/plot_smoke.R")
source("code/time_series.R")
source("code/create_time_covid_plots.R")
```

I want to compare how Covid affected African countries relative to other continents. In order to do this I'm going to compare various COVID-19 statistics over 6 month intervals for every continent. Therefore I will divide the times period into 5 periods: 2020-01, 2020-07, 2021-01, 2021-07 and 2022-01. I will give the mean value for each of these continents over each 6 month period.

In order to do this I need to build a function which calculates the mean value for a particular country over this a particular 6 month period. 

I want to look at the mean value for total_cases_by_million for each country so I'll use the function calculate the mean for each country based on each period.

```{r}
library(tidyverse)

coviddf <- data_transformation(coviddf)
```

Ok so this function is giving me errors and is not working so I'm going to try simplify it to calculate the mean_total cases for each country in that particular period, first I'm going to add a period column to the main data frame indicating the 5 different periods.

Now I generate the bargraphs showings the growing rates of the average cases, hospitalisations, deaths and vaccinations.

```{r}

# Generate mean cases per million plots
covid_cases <- create_covid_plots(coviddf, "mean_cases", "Mean Cases Per Million", "continent", "Period", "period")
covid_cases

# Generate mean daily hospitilizations per million plots
covid_hosp <- create_covid_plots(coviddf, "mean_hosp", "Mean Daily Hospitilizations Per Million", "continent", "Period", "period")
covid_hosp

# Generate mean deaths per million plots
covid_deaths <- create_covid_plots(coviddf, "mean_deaths", "Mean Deaths Per Million", "continent", "Period", "period")
covid_deaths

# Generate mean vaccinations per million plots
covid_vacs <- create_covid_plots(coviddf, "mean_vacs", "Mean Vaccinations Per Million", "continent", "Period", "period")
covid_vacs

```


Ok so as can be seen in the results Africa is the least affected by Covid on all accounts. Of all the continents it has the lowest number of cases, hospitalisations, deaths and vaccinations per million of every continent. It appears the more developed continents, Europe and North America are the most significantly affected.

For the next part of my analysis I want to look at the correlation between smoking and hospitalisations to see if countries with an higher rate of smoking are more likely to be hospitalised. I need to import the death_by_cause data to look at smoking related deaths.

```{r}
deathdf <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_1\\data\\Covid\\Deaths_by_cause.csv")
```

Because Europe was most significantly affected by Covid I want to only look at European countries to do my comparison, I therefore want to look at those countries which had the greatest level of lower respiratory infections as these can cause higher death rates amongst smokers as they have weaker lungs. Therefore I need to look at the European countries with the 3 greatest number of average daily hospitalisations and the 3 lowest. I will find the hospitialisations and populations of all measured European countries.

```{r}

Eur_hosp_smoke <- smoke_data_trans(coviddf)

Eur_hosp_smoke
```

Ok so based on this it shows that the highest daily hospitalisations for Covid in Europe over the five periods were in Bulgaria, Serbia, Lithuania, Romania and Hungary whilst the lowest recorded daily hospitalisations were in Norway, Finland, Iceland, Denmark and the Netherlands. 

Now I need to look at each countries prevalence for smoking by averaging their respiratory infection deaths form the Deaths_by_cause data set.

```{r}
avg_resp_deaths <- resp_trans(deathdf)
```

Now I need to combine these two sets to compare respiratory deaths with covid hospitalisation and create a respiratoty deaths per population variable.

```{r}
Eur_hosp_smoke <- merge_resp_hosp(Eur_hosp_smoke)
```

Now that I have the average number of Covid hospitalisations and the number of lower respiratory infection deaths per million of the population I can create a scatter polt to see if their was any correlation.

```{r}

scatter_resp <- plot_smoke(Eur_hosp_smoke)
scatter_resp


resp_death_table <- Eur_hosp_smoke %>%
  select(location, mean_hosp, percent_resp_deaths) %>% 
    tibble()

resp_death_table
```

So as can be seen there is almost no correlation between the historical respiratory deaths and daily Covid hospitilations for countries in Europe. This would suggest that countries with historically high levels of smoking are unlikely to have more or less covid-related hospitalisations. 

For the final part of the question I'm going to look at hospitalisation numbers and ICU numbers for each continent and see if the one led or lagged the other. Ok so first I'm going to group the data by continent and by month.

```{r}
library(lubridate)

time_covid <- time_series(coviddf)
time_covid
```

Now I want to plot the data to see if hospitalisations number lead or lag ICU numbers.

```{r}
# Generate plots for Europe
Eur_ICU <- create_time_covid_plots(time_covid, "Europe")
Eur_ICU

# Generate plots for Asia
Asia_ICU <- create_time_covid_plots(time_covid, "Asia")
Asia_ICU

# Generate plots for North America
NA_ICU <- create_time_covid_plots(time_covid, "North America")
NA_ICU

```

Therefore it appears that the level of hospitalizations, which is being used as a proxy for hospital capacity, coincides with an increase in ICU admissions for Europe, Asia and North America when looking at the monthly admissions.

