---
title: "Question 4"
documentclass: elsarticle
Thesis_FP: no
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: yes
    template: Tex/TexDefault.txt
    fig_width: 3.5
    fig_height: 3.5
Author1: Peter Meihuizen
Ref1: Stellenbosch University
Email1: 21831041\sun.ac.za
BottomRFooter: \footnotesize Page \thepage
addtoprule: yes
addfootrule: yes
margin: 2.3
bottom: 2
top: 2.5
HardSet_layout: yes
linenumbers: no
bibliography: Tex/ref.bib
csl: "Tex/harvard-stellenbosch-university.csl"
RemovePreprintSubmittedTo: yes
toc: no
numbersections: yes
fontsize: 11pt
linestretch: 1.2
link-citations: yes
AddTitle: yes
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

Ok so lets see what is popular on netflix.

# Analysis \label{Analysis}

Let's first import the data

```{r}
netflix <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_4\\data\\netflix\\titles.csv")
source("code/top10.R")
source("code/top200.R")
```

We now have a netflix data set showing the information on shows and movies. This also shows the rating an votes made on IMDB, a well respected website which gives reliable ratings while accurately reflecting public opinion towards shows and movies. Therefore I will use IMDB scores as a metric for quality of a show or movie, and I will use the number of IMDB votes as a metric for the number of people who watched the show/movie. To see if the more frequently watched movies are also highly rated I will plot the 10 most voted movies and there IMDB scores.

```{r}
top_10_imdb_votes <- netflix %>% 
  top10(imdb_votes, title, type, release_year, imdb_score, production_countries, x = 10)

top_10_imdb_votes
```

As can be seen the most voted movies and shows on IMDB all also have very high ratings. They are also very well_known movies and tv shows. This shows that well-known and highly rated movies/shows are the most popular amongst viewers. It can also be seen that all the top picks are made by US or GB production companies. 

In order to be a successful streaming company, we need people to watch our content therefore I will use imdb_votes as the metric for viewership and see what factors attracts people to watch particular content. Using the netflix data set, I am  going to look at the top 200 shows most viewed shows and movies and see the breakdown of these shows and movies by genre.

Ok so I want to build a function which shows the distribution of the top 200 shows on netflix. Ok so I'm having issues with my function so I'm just going to create two data sets each with only movies and only shows in case I want to see the most popular movie genres also.

```{r}
movies <- filter(netflix, grepl("MOVIE", netflix$type)) 
shows <- filter(netflix, grepl("SHOW", netflix$type))
```

Ok now that I have seperated the data sets, we can work out the most popular 200 shows to see what genres are popular.  

```{r}
top200_shows <- top200(shows, "Genres of Top 200 Shows")
top200_shows

top200_movies <- top200(movies, "Genres of Top 200 Movies")
top200_movies
```

Ok so as can be seen by both bargraphs, the top shpws and movies which are watched are drama and comedy. Drama is the most voted by quite a long way for both, but comedy also is significantly more popular than the remaining genres. All of the remaining genres are pretty similar in popularity. This shows that if you want to make a successful streaming platform, your content should be predominantly made up of drama shows and movies, as well as comedies.
