---
title: "Question 5"
documentclass: elsarticle
Thesis_FP: no
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: yes
    template: Tex/TexDefault.txt
    fig_width: 3.5
    fig_height: 3.5
Author1: Peter Meihuizen
Ref1: Stellenbosch University
Email1: 21831041\@sun.ac.za
keywords: Multivariate GARCH \sep Kalman Filter \sep Copula
JELCodes: L250 \sep L100
BottomRFooter: \footnotesize Page \thepage
addtoprule: yes
addfootrule: yes
margin: 2.3
bottom: 2
top: 2.5
HardSet_layout: yes
linenumbers: no
bibliography: Tex/ref.bib
csl: "Tex/harvard-stellenbosch-university.csl"
RemovePreprintSubmittedTo: yes
toc: no
numbersections: yes
fontsize: 11pt
linestretch: 1.2
link-citations: yes
AddTitle: yes
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

For this question we will be looking at the database of app downloads from Google to see what encourages people to download a particular app.

# Analysis \label{Analysis}

Firstly I need to import the data.

```{r}
Google <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_5\\data\\googleplay\\googleplaystore.csv")
source("code/scatter.R")
source("code/Review_table.R")
```

Ok so based on the data I want to see what are the most downloaded types of apps, therefore I will look at the number of installations per application type. First I need to change this line to be numeric values.

```{r}
Google$Installs <- as.numeric(gsub("[+,]", "", Google$Installs))
```

Ok so now I want to see which app categories are the most used. In order to do that I am going to do a scatter plot which shows the average number of reviews of each app category, as well as the average number of downloads received by each category.

```{r}
Review_Install_scatter <- scatterplot(Google)
Review_Install_scatter
```

Ok so we can see that communication apps are by far the most installed type of app on average, having the highest average number of reviews and installations. In terms of installations it is followed by social apps and then gaming apps. The second most reviewed category seems to be video_players.

So we know what is the most used types of apps and most reviewed. However an important factor when understanding the reviews of a particular is app is whether or not they are positive. Apps with more positive reviews are likely to be used more frequently, so using the user review data set, we can compare the percentage of positive, negative and neutral reviews for each app category.

```{r}
Review <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_5\\data\\googleplay\\googleplaystore_user_reviews.csv")
```

Ok now we must merge the category column to be assigned with therespective app in the Review data set.

```{r}
category <- Google %>% select(App, Category)

merged_data <- Review %>%
  left_join(category, by = "App")
```

Ok so now using the merged data we can work out the percentage of each type of review in each category.

```{r}
Pos_neg_reviews <- Review_table(merged_data)

Pos_neg_reviews
```

So as can be seen most app categories have round 60% positive reviews, 25% negative reviews and 15% neutral reviews. Communication apps have a good positive percentage relative to other app types. They have the best percentage results amongst their closest competitors in terms of installations, with 63% positive reviews and 21% negative reviews. Comparatively, social apps having 54% positive reviews and 26% negative reviews, whilst gaming apps have 58% positive reviews and 37% negative reviews. Based on these results it seems evident that the best type of app to produce would be a Communication app.


