---
title: "Question 2"
documentclass: elsarticle
Thesis_FP: no
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: yes
    template: Tex/TexDefault.txt
    fig_width: 3.5
    fig_height: 3.5
Author1: Peter Meihuizen
Ref1: Stellenbosch Univerity, South Africa
Email1: 21831041\@sun.az.za
BottomRFooter: \footnotesize Page \thepage
addtoprule: yes
addfootrule: yes
margin: 2.3
bottom: 2
top: 2.5
HardSet_layout: yes
linenumbers: no
bibliography: Tex/ref.bib
csl: "Tex/harvard-stellenbosch-university.csl"
RemovePreprintSubmittedTo: yes
Journal: Journal of Finance
toc: no
numbersections: yes
fontsize: 11pt
linestretch: 1.2
link-citations: yes
AddTitle: yes
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

For this question my goal is to prove to my friend that the weather in the UK is terrible.

# Analysis \label{Analysis}

To conduct my analysis I must first import my data files.

```{r}
london <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_2\\data\\London\\london_weather.csv")
source("code/plot_midlands.R")
source("code/midlands_transform.R")
source("code/london_weather.R")
source("code/plot_london.R")
source("code/transform_data.R")
```

Ok so first I want to plot a temperature graph showing the mean temperature for each month for both london and the midlands datasets. I will make a function which does this and therefore I can replicate the process with other weather elements. I'm interested in convincing you not to move there, therefore I am only concerned with assessing the weather since the turn of the 21st century. 

```{r}
library(lubridate)

london <- transform_data(london)
```


Plotting a bargraph of the average temperature, sunshine, rain and cloud cover in London per month.


```{r}
# Generate Average Monthly Temperature plot for London
temp_lon <- london_weather(london, quote(avg_temp_month), "Average Temperature",
                            "Average Monthly Temperature for London", "pink")
temp_lon

# Generate Average Monthly Precipitation plot for London
prec_lon <- london_weather(london, quote(avg_prec_month), "Average Precipitation",
                            "Average Monthly Precipitation for London", "blue")
prec_lon

# Generate Average Monthly Sunshine plot for London
sun_lon <- london_weather(london, quote(avg_sun_month), "Average Sunshine",
                            "Average Monthly Sunshine for London", "yellow")
sun_lon

# Generate Average Monthly Cloud Cover plot for London
clo_lon <- london_weather(london, quote(avg_cloud_month), "Average Cloud Cover",
                            "Average Monthly Cloud Cover for London", "grey")
clo_lon
```

This measures the average for each month over a 20 year period and as can be seen the weather in London is pretty miserable. In their best months they only get 6 hours of sun on average a day, even though it is daytime for over 16 hours in those months. They get constant rain and cloud cover throughout the year and their average temperature doesn't even get above 20 in their best months. But you want to live in the mid parts of england so lets see how these stats look compared to London.

Now we must import the data for the midlands in the UK and look at the monthly weather here.

```{r}
midlands <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_2\\data\\London\\UKMonthly_Detailed.csv")
```

Now we transfrom the data.

```{r}
library(dplyr)
library(lubridate)

midlands <- midlands_transform(midlands)
```

Ok so the only values we have are for temperature and precipitation. Once again we're looking at the average temperature for each month over a 20 year period. However for precipitation the data shows the amount of rain on the most rainy day in the month. These values are then average over 20 years.

```{r}
temp_mid <- ggplot(midlands, aes(x = month, y = avg_temp_month)) +
  stat_summary(geom = "bar", fill = "pink", fun = "mean") +
  labs(x = "Month", y = "Average Temperature", title = "Average Monthly Temperature for Midlands") +
  theme_bw()

temp_mid

prec_mid <- ggplot(midlands, aes(x = month, y = avg_prec_month)) +
  stat_summary(geom = "bar", fill = "blue", fun = "mean") +
  labs(x = "Month", y = "Average Precipitation", title = "Average Monthly Precipitation for Midlands") +
  theme_bw()

prec_mid
```

As you can see it is even colder in the midlands where you plan to go, every month is nearly 5 degrees colder than in London. Also the most rainy days of the year happen in Summer in the midlands based on these graphs. Even their warmest months have lots of rain. Therefore it is pretty evident that the weather in the UK is, as expected, pretty terrible and you should definitely rather stay in Cape Town.

