---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 3"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Peter Meihuizen"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University" # First Author's Affiliation
Email1: "21831041\\@sun.ac.za" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
#  Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
Example_data <- Texevier::Ex_Dat

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.
write_rds(Example_data, path = "data/Example_data.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

This question shows the popularity of different songs for Metallica and Coldplay. I will try to analyse these different songs to see what makes a song popular.

# Analysis \label{Analysis}

I first need to import the data for both Coldplay and Metallica's songs.

```{r}
Coldplay <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_3\\data\\Coldplay_vs_Metallica\\Coldplay.csv")

Metallica <- read.csv("C:\\Users\\Peter Meihuizen\\Desktop\\Masters\\1st Semester\\Data Science\\Take-home test\\Question_3\\data\\Coldplay_vs_Metallica\\metallica.csv")
source("code/create_scatterplot.R")
source("code/boxplot_create.R")
```

So I want to look at what makes songs popular, so I'm going to do some scatter plots to plot the songs various aspects with their popularity to see how they compare. To do this I'm going to build a function which makes a scatterplot.

Plotting the scatterplot shows as follows:

```{r}
Coldplay %>% 
    create_scatterplot(valence, popularity, title = "Coldplay Songs Popularity and Positivity")
```

It can be seen that Coldplay's music mostly has a lower valence, meaning it resembles mostly negative emotions, and it appears that there is a slight negative correlation between vlence and popualrity, suggesting more positive music does slightly worse than more negative music.

Everyone likes to dance right? So what effect does the danceability have on the popularity of their songs?

```{r}
Coldplay %>% 
    create_scatterplot(danceability, popularity, title = "Coldplay Songs Popularity and Tempo")
```

Again it appears that danceability has a positive effect on the popularity of Coldplay's songs.

Ok what about Metallica?

```{r}
Metallica %>% 
    create_scatterplot(valence, popularity, title = "Metallica Songs Popularity and Positivity")
```

There seems to be minimal correlation between valence and popularity for Metallica's songs. It seems as though their songs are more concentrated around a low valence, suggesting they express negative emotions. However their more positive songs are slgihtly more popular as is suggested by the very gradual slope of the correlation line.

```{r}
Metallica %>% 
    create_scatterplot(danceability, popularity, title = "Metallica Songs Popularity and Tempo")
```

Again, the same as was seen with Coldplay, danceability increases the popularity of Metallica's songs, despite the majority of their songs being lower on the danceability scale.

Ok so it seems as though danceability makes a particular bands songs more popular. 
So let's see how the popularity and danceablility was distributed across their various albums and see whether these are linked. In order to do this lets make a function that looks at the distribution of the danceability of the different albums of Coldplay and Metallica.

I'm just going to rename the album_name column in the Coldplay data frame so that it is the same as in the Metallica data set. I am also going to get rid of the live albums and songs.

```{r}
Coldplay <- Coldplay %>%
    rename(album = album_name) %>%
    filter(!grepl("live|Live", album)) %>% 
    filter(!grepl("live|Live", name))
```

Now ploting the boxplots, first by popularity:

```{r}
Coldplay_pop <- Coldplay %>% 
    boxplot_create(popularity, y_lab = "Popularity", title = "Distribution of Popularity by Coldplay Album")

Coldplay_pop
```

So as can be seen Coldplay's albums vary quite a lot in terms of there popularity, some were significantly more popular than others. What abou the daceability of their albums?

```{r}
Coldplay_dance <- Coldplay %>% 
    boxplot_create(danceability, y_lab = "Danceability", title = "Distribution of Danceability by Coldplay Album")

Coldplay_dance
```

It appears as though the distribution of danceable songs on Coldplay's albums have been pretty similar across albums. 

How about Metallica now. First I need to remove the live songs and albums.

```{r}
Metallica <- Metallica %>%
    filter(!grepl("live|Live", album)) %>% 
    filter(!grepl("live|Live", name))
```

Now to look at the popularity and danceability of the albums. First popularity.

```{r}
Metallica %>% 
    boxplot_create(popularity, y_lab = "Popularity", title = "Distribution of Popularity by Metallica Album")
```

Ok similar to Coldplay there are some differences in popularity between Matallica's albums. Now let's look at their danceability.

```{r}
Metallica %>% 
    boxplot_create(danceability, y_lab = "Danceability", title = "Distribution of Danceability by Metallica Album")
```

As can be seen there's more variation within albums around the dancability of Metallica's songs, however this variation is seen in most of there albums suggesting they do not really change their methods between albums.

